---
title: "Output of multivariate analysis"
author: "Victor Ordu"
date: "22 May 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## This is a work in progress

**The source file for this analysis is at multivariate-analysis.R**
```{r, echo=FALSE}
library(MASS)
library(Hmisc)
library(tidyr)


data<- readRDS("clean_x.rds")
mydata <- data[, c(2, 5, 6, 12, 13, 8)] # extract variables of interest
rm(data)

# Partition the dataset
index <- caret::createDataPartition(mydata$concern.stress, times = 1, p = .8,
                                    list = FALSE)
train <- mydata[index, ]
test <- mydata[-index, ]
rm(index)

# Build array and multi-way contingency table
covars <- ~ risks.time + risks.communication + risks.lackemployeecontrol + 
             risks.unclearPolicy + risks.workinghours + concern.stress
array <- xtabs(covars, data = train); rm(covars)
array
print(ftable(array))
```

Visualize this arrangement
```{r, echo=FALSE}
vcd::mosaic(array)
```

```{r, echo=FALSE}
vcd::assoc(array)
rm(array, covars)
```
**Proportional odds logistic regression**
# Prep: Define the formula for modelling
```{r, echo=FALSE}
orgFormula <- concern.stress ~ risks.time + risks.communication + risks.lackemployeecontrol + 
  risks.unclearPolicy +  risks.workinghours         # organisational issues
```
```{r}
ordered.fit <- polr(formula = orgFormula, data = mydata, Hess = TRUE)
summary(ordered.fit)

# Tabulate coefficients 
table.coef <- coef(summary(ordered.fit))
table.coef

# calculate and store p-values
pval <- pnorm(abs(table.coef[, "t value"]), lower.tail = FALSE) * 2
table.coef <- cbind(table.coef, "p-value" = pval)
table.coef
rm(table.coef, pval)
```

Calculate 95% confidence intervals
```{r}
CI <- confint(ordered.fit)
CI
confint.default(ordered.fit)
```
Obtain the Odds Ratio and match with confidence intervals
```{r}
odd.ratio <- exp(coef(ordered.fit))
odd.ratio

# Tabulate O.R. and CI
(odds.tabl <- cbind("O.R." = odd.ratio, exp(CI)))
```
```{r} rm(odd.ratio, CI, odds.tabl)```


Assessing the model
```{r}
# --- Evaluating the assumption of proportional odds ---
# A function to plot at class cut-off points
props <- function(y) {
  c('Y >= 1' = qlogis(mean(y >= 1)),
    'Y >= 2' = qlogis(mean(y >= 2)),
    'Y >= 3' = qlogis(mean(y >= 3)))
}

trial <- with(mydata,
               summary(as.numeric(concern.stress) ~ risks.time + 
                         risks.communication + risks.lackemployeecontrol + 
                         risks.unclearPolicy + risks.workinghours, fun = props))
trial
```

```{r}
# Treat outcome as a binary variable via "cut-off" points to test assumption
glm(I(as.numeric(concern.stress) >= 2) ~ risks.time + 
      risks.communication + risks.lackemployeecontrol + 
      risks.unclearPolicy + risks.workinghours, family = "binomial",
    data = mydata)

glm(I(as.numeric(concern.stress) >= 3) ~ risks.time + 
      risks.communication + risks.lackemployeecontrol + 
      risks.unclearPolicy + risks.workinghours, family = "binomial",
    data = mydata)
```

```{r}
trial[, 4] <- trial[, 4] - trial[, 3]
trial[, 3] <- trial[, 3] - trial[, 3]
trial
```
Plot the logits
```{r}
plot(trial, which = 1:3, pch = 1:3, xlab = "logit", main = "",
     xlim = range(trial[, 3:4]))
```
```{r}
rm(trial)
```

Fit the model if assumption is met
```{r}
# Create a dummy dataframe to predict probabilities
kays <- rep(c("Yes", "No"), 600)
data.pred <- data.frame(risks.time = kays,
                        risks.communication = kays,
                        risks.lackemployeecontrol = kays,
                        risks.unclearPolicy = kays,
                        risks.workinghours = kays)
data.pred <- cbind(data.pred, predict(ordered.fit, data.pred, type = "probs"))
head(data.pred)
```
```{r}
# Tidy the dataframe
data.pred <- data.pred %>%
  gather(category, probability, `No concern`:`Major concern`,
                factor_key = TRUE)
head(data.pred)
```

```{r, echo=FALSE}
# cleaning up search path and the workspace
Vectorize(detach)(name = paste0("package:",c("MASS", "Hmisc", "lattice", "survival", "Formula", "ggplot2", "tidyr")), character.only = TRUE)
rm(list = ls())
```
**End**
